---
title: "U.S. 2024 Election Results: A National Vibe Shift"
format:
  html:
    toc: true
    toc-location: right
    toc-title: "On this page"
    css: styles.css
---
```{r}
#| code-fold: true
#| warning: false
#| message: false
#| results: 'hide'
# Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
library(jsonlite)
library(purrr)
library(httr2 )
library(rvest)
library(janitor)
library(sf)
library(scales)
```
::: {.justify style="text-align: justify;"}
# Executive Summary
This project explores changes in U.S. county-level voting behavior between the 2020 and 2024 Presidential elections. Election results were scraped from Wikipedia due to inconsistencies in official state election websites and official county boundary shapefiles were obtained from the U.S. Census Bureau. Key findings from the analysis include both absolute and proportional changes in voting patterns, turnout and geographic characteristics. Some counties recorded large gains in Republican votes while others showed strong Democratic support, particularly in smaller or urban areas. The results highlight the complexity of interpreting nationwide electoral trends based solely on visual maps.
:::

::: {.justify style="text-align: justify;"}
# Highlights of Key Findings
- *Largest Number of Trump Votes (2024)*:
Los Angeles County, California cast the highest number of votes for Donald Trump in 2024, over 1.18 million votes. However, this may be due to the county's large population, not a Republican majority.
- *Highest Proportion of Biden Votes (2020)*:
Kalawao County, Hawaii recorded 95.8% of its votes for Joe Biden, the highest percentage in the nation, though from a very small voter base.
- *Largest Absolute Gain for Trump (2024)*:
Los Angeles County, California also saw the largest increase in Trump votes from 2020 to 2024, an increase of 44,332 votes.
- *Biggest Shift Toward Harris (2024)*:
Florida experienced the largest shift in vote margin in favor of Kamala Harris with a gain of approximately 371,686 votes.
- *Largest County by Land Area*:
Yukon-Koyukuk County, Alaska is the largest in the dataset, covering over 377 billion square meters, aligning with known geographic facts.
- *Highest Voter Density (2020)*:
New York County, New York had the highest voter density, with 0.01 voters per square meter, reflecting high population density in a small area.
- *Highest Increase in Voter Turnout (2020–2024)*:
Montgomery County, Texas saw the biggest rise in total votes cast, increasing by 35,715 votes from 2020 to 2024.
:::

# Introduction
::: {.justify style="text-align: justify;"}
After the 2024 U.S. Presidential election, The New York Times published a map suggesting that many counties across the United States had shifted toward the Republican party compared to 2020. This visual quickly spread online and shaped public discussions around a possible nationwide political “vibe shift.” This project aims to investigate the changes in vote shares between the 2020 and 2024 presidential elections at the county level as well as recreate and evaluate the original map. 
Data on election results at the county-level was collected for both years from Wikipedia pages for each state. While not a primary source, Wikipedia provided a reliable and practical alternative due to the challenges of collecting data from official state websites, which are often inconsistent and difficult to automate. For mapping, the U.S. county boundary shapefiles from the U.S. Census Bureau’s Cartographic Boundary Files (2023) was used and the finest available resolution (1:500,000) was selected to ensure high-quality geographic visualizations.
By combining cleaned election results with accurate county shapes, the shift in Republican vote share in each county was calculated and visualized using tools in R. This project offers a data-driven look at whether the supposed nationwide political shift is supported by the numbers and where that shift, if any, was most pronounced.
:::

# Data Aquisition
## Task 1: US County Shapefiles
::: {.justify}
The shapefile containing U.S. county (and equivalent) boundaries was obtained from the U.S. Census Bureau’s Cartographic Boundary Files (2023).The finest available resolution (1:500,000) was chosen to ensure high-quality and accurate geographic visualizations.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| results: 'hide'
# Create directory if it doesn't exist
if (!file.exists("data/mp04")) {
  dir.create("data/mp04", showWarnings = FALSE, recursive = TRUE)
}

# Define the URL and destination file path (2023, 1:500k resolution)
url <- "https://www2.census.gov/geo/tiger/GENZ2023/shp/cb_2023_us_county_500k.zip"
zip_path <- "data/mp04/cb_2023_us_county_500k.zip"

#Download the file if it's not already downloaded
if (!file.exists(zip_path)) {
  download.file(url, destfile = zip_path, mode = "wb")
  message("2023 Shapefile downloaded.")
} else {
  message("2023 Shapefile already downloaded.")
}

# Unzip the file if not already unzipped
shp_file_exists <- any(grepl("\\.shp$", list.files("data/mp04")))
if (!shp_file_exists) {
  unzip(zip_path, exdir = "data/mp04")
  message("2023 Shapefile unzipped.")
} else {
  message("2023 Shapefile already unzipped.")
}
```

::: {.justify style="text-align: justify;"}
The 2020 and 2024 U.S. Presidential election results at the county level were collected from Wikipedia pages for each state. Since official state websites are not consistent and difficult to automate, Wikipedia was used as a reliable alternative. Using R packages like httr2 and rvest, the tables with county-level results were downloaded, cleaned, and labeled with the state name to prepare the data for analysis
:::

## Task 2: Acquire 2024 US Presidential Election Results
```{r}
#| code-fold: true
#| warning: false
#| message: false
#| results: 'hide'
#Create directory to store downloaded HTML pages
if (!dir.exists("data/mp04/wikipedia_pages")) {
  dir.create("data/mp04/wikipedia_pages", recursive = TRUE)
}

# List of 50 U.S. States
us_states <- c(
  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut",
  "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa",
  "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan",
  "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
  "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio",
  "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
  "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia",
  "Wisconsin", "Wyoming"
)

#  Safe filename helper function
safe_filename <- function(state) {
  gsub(" ", "_", tolower(state))
}

# Scraping function for a single state
scrape_state_results <- function(state) {
  message("Processing: ", state)
  
  # Construct Wikipedia path and local cache path
  base_url <- "https://en.wikipedia.org/wiki"
  path <- paste0("2024_United_States_presidential_election_in_", gsub(" ", "_", state))
  html_path <- file.path("data/mp04/wikipedia_pages", paste0(safe_filename(state), ".html"))
  
  # Download HTML only if not already saved
  if (!file.exists(html_path)) {
    req <- request(base_url) |>
      req_url_path_append(path) |>
      req_perform()
    writeBin(resp_body_raw(req), html_path)
  }
  
  # Read the HTML and extract all tables
  html <- read_html(html_path)
  tables <- html_elements(html, "table") |> html_table(fill = TRUE)
  
  # Find table with a column like "County", "Parish", "Borough", etc.
  target_table <- NULL
  for (tbl in tables) {
    colnames_lower <- tolower(names(tbl))
    if (any(str_detect(colnames_lower, "county|parish|borough|city and borough"))) {
      target_table <- tbl
      break
    }
  }
  
  # If not found, return NULL with warning
  if (is.null(target_table)) {
    warning("No county-level table found for ", state)
    return(NULL)
  }
  
  # Clean and return table with state column
  cleaned <- target_table |>
    clean_names() |>
    mutate(state = state)
  
  return(cleaned)
}

# Scrape all 50 states 
all_election_results <- map_dfr(us_states, possibly(scrape_state_results, otherwise = NULL))

#Save final result
write_csv(all_election_results, "data/mp04/all_county_election_results_2024.csv")

# Preview the combined dataset
glimpse(all_election_results)
```
## Task 3: Acquire 2020 US Presidential Election Results
```{r}
#| code-fold: true
#| warning: false
#| message: false
#| results: 'hide'
# Create directory to store downloaded 2020 pages
if (!dir.exists("data/mp04/wikipedia_pages_2020")) {
  dir.create("data/mp04/wikipedia_pages_2020", recursive = TRUE)
}

# List of 50 U.S. states
us_states <- c(
  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut",
  "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa",
  "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan",
  "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire",
  "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio",
  "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
  "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia",
  "Wisconsin", "Wyoming"
)

# Safe filename helper
safe_filename <- function(state) {
  gsub(" ", "_", tolower(state))
}

# Scraping function for 2020 results
scrape_2020_results <- function(state) {
  message("Processing: ", state)
  
  base_url <- "https://en.wikipedia.org/wiki"
  path <- paste0("2020_United_States_presidential_election_in_", gsub(" ", "_", state))
  html_path <- file.path("data/mp04/wikipedia_pages_2020", paste0(safe_filename(state), ".html"))
  
  # Download if not cached
  if (!file.exists(html_path)) {
    req <- request(base_url) |>
      req_url_path_append(path) |>
      req_perform()
    writeBin(resp_body_raw(req), html_path)
  }
  
  # Read and parse HTML
  html <- read_html(html_path)
  tables <- html_elements(html, "table") |> html_table(fill = TRUE)
  
  # Find relevant table
  target_table <- NULL
  for (tbl in tables) {
    colnames_lower <- tolower(names(tbl))
    if (any(str_detect(colnames_lower, "county|parish|borough|city and borough"))) {
      target_table <- tbl
      break
    }
  }
  
  if (is.null(target_table)) {
    warning("No county-level table found for ", state)
    return(NULL)
  }
  
  cleaned <- target_table |>
    clean_names() |>
    mutate(state = state)
  
  return(cleaned)
}

# Scrape all 50 states
all_results_2020 <- map_dfr(us_states, possibly(scrape_2020_results, otherwise = NULL))

# Save results
write_csv(all_results_2020, "data/mp04/all_county_election_results_2020.csv")

# Preview
glimpse(all_results_2020)
```

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| results: 'hide'
#Combining all data sets
# Load county shapefile
county_shapes <- st_read("data/mp04/cb_2023_us_county_500k.shp") |>
  clean_names()

# Load 2020 election results
results_2020 <- read_csv("data/mp04/all_county_election_results_2020.csv") |>
  clean_names()

# Load 2024 election results
results_2024 <- read_csv("data/mp04/all_county_election_results_2024.csv") |>
  clean_names()

# Standardize names for merging
results_2020 <- results_2020 |>
  mutate(county = str_to_lower(str_remove_all(county, " county| parish| borough")),
         state = str_to_lower(state))

results_2024 <- results_2024 |>
  mutate(county = str_to_lower(str_remove_all(county, " county| parish| borough")),
         state = str_to_lower(state))

county_shapes <- county_shapes |>
  mutate(county = str_to_lower(str_remove_all(name, " county| parish| borough")),
         state = str_to_lower(state_name))

# Join election results to spatial data
shapes_2020 <- left_join(county_shapes, results_2020, by = c("state", "county"))
shapes_2024 <- left_join(county_shapes, results_2024, by = c("state", "county"))

# Combine 2020 and 2024 data
elections_combined <- full_join(
  results_2020 |> rename_with(~paste0(.x, "_2020"), -c(county, state)),
  results_2024 |> rename_with(~paste0(.x, "_2024"), -c(county, state)),
  by = c("county", "state")
)

# Join combined results with shapefile
shapes_combined <- left_join(county_shapes, elections_combined, by = c("county", "state"))

# Preview the result
glimpse(shapes_combined)
```
## Task 4: Initial Analysis Questions

### 1.Which county or counties cast the most votes for Trump (in absolute terms) in 2024?
::: {.justify style="text-align: justify;"}
The result indicates that in the 2024 U.S. presidential election, Los Angeles County in California cast the highest number of votes for Donald Trump, with a total of 1,189,862 votes. This large figure reflects the county's massive population, not necessarily strong Republican dominance. Despite being a traditionally Democratic-leaning area, Los Angeles’ sheer size means even minority support can result in high absolute vote counts.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Convert Trump vote count column to numeric if it's not already
# (Skip this if it's already numeric, just here for robustness)
if (!is.numeric(shapes_combined$donald_trump_republican_2024)) {
  shapes_combined <- shapes_combined |>
    mutate(donald_trump_republican_2024 = readr::parse_number(donald_trump_republican_2024))
}

# Find the county or counties with the highest Trump vote count in 2024
top_trump_counties_2024 <- shapes_combined |>
  filter(!is.na(donald_trump_republican_2024)) |>
  arrange(desc(donald_trump_republican_2024)) |>
  slice_max(donald_trump_republican_2024, n = 1) |>
  select(state, county, donald_trump_republican_2024, geometry)

# Display the result in a clean kable table
top_trump_counties_2024 |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    donald_trump_republican_2024 = comma(donald_trump_republican_2024)
  ) |>
  kable(
    caption = "County with the Most Votes for Donald Trump in 2024",
    col.names = c("State", "County", "Trump Votes (2024)"),
    align = "lcl",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 2.Which county or counties cast the most votes for Biden (as a fraction of total votes cast) in 2020?

::: {.justify style="text-align: justify;"}
The result shows that in the 2020 U.S. presidential election, Kalawao County in Hawaii cast the highest proportion of its total votes for Joe Biden with 95.8% support. Although this percentage is the highest nationally, it reflects results from a very small voting population. This highlights how smaller counties can show extremely high vote shares even when the absolute number of votes is low.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
#Convert columns to numeric if needed
if (!is.numeric(shapes_combined$joe_biden_democratic_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(joe_biden_democratic_2020 = parse_number(joe_biden_democratic_2020))
}

if (!is.numeric(shapes_combined$total_votes_cast_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(total_votes_cast_2020 = parse_number(total_votes_cast_2020))
}

#Compute Biden's share of total votes
shapes_combined <- shapes_combined |>
  mutate(biden_vote_percentage_2020 = (joe_biden_democratic_2020 / total_votes_cast_2020) * 100)

#Identify county or counties with the highest Biden vote share
top_biden_counties_2020 <- shapes_combined |>
  filter(!is.na(biden_vote_percentage_2020)) |>
  arrange(desc(biden_vote_percentage_2020)) |>
  slice_max(biden_vote_percentage_2020, n = 1) |>
  select(state, county, biden_vote_percentage_2020, geometry)

#Display results in a formatted table
top_biden_counties_2020 |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    biden_vote_percentage_2020 = percent(biden_vote_percentage_2020 / 100, accuracy = 0.1)
  ) |>
  kable(
    caption = "County with the Highest Percentage of Votes for Joe Biden in 2020",
    col.names = c("State", "County", "Biden Vote Percentage (2020)"),
    align = "lcl",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 3.Which county or counties had the largest shift towards Trump (in absolute terms) in 2024?

::: {.justify style="text-align: justify;"}
In the 2024 U.S. Presidential Election, Los Angeles County in California recorded the largest increase in votes for Donald Trump compared to 2020. Trump received 44,332 more votes in 2024 than he did in 2020, rising from 1,145,530 votes to 1,189,862 votes. This shift indicates a notable gain in support for Trump in the nation’s most populous county.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Ensure both Trump vote columns are numeric
if (!is.numeric(shapes_combined$donald_trump_republican_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(donald_trump_republican_2020 = parse_number(donald_trump_republican_2020))
}

if (!is.numeric(shapes_combined$donald_trump_republican_2024)) {
  shapes_combined <- shapes_combined |>
    mutate(donald_trump_republican_2024 = parse_number(donald_trump_republican_2024))
}

# Calculate absolute vote gain for Trump
shapes_combined <- shapes_combined |>
  mutate(shift_to_trump = donald_trump_republican_2024 - donald_trump_republican_2020)

# Find the county or counties with the largest gain
largest_trump_shift <- shapes_combined |>
  filter(!is.na(shift_to_trump)) |>
  arrange(desc(shift_to_trump)) |>
  slice_max(shift_to_trump, n = 1) |>
  select(state, county, donald_trump_republican_2020, donald_trump_republican_2024, shift_to_trump)

# Display final result in clean HTML table
largest_trump_shift |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    donald_trump_republican_2020 = comma(donald_trump_republican_2020),
    donald_trump_republican_2024 = comma(donald_trump_republican_2024),
    shift_to_trump = comma(shift_to_trump)
  ) |>
  kable(
    caption = "County with the Largest Shift towards Trump 2024",
    col.names = c("State", "County", "Trump Votes 2020", "Trump Votes 2024", "Vote Gain"),
    align = "lclrr",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 4.Which state had the largest shift towards Harris (or smallest shift towards Trump) in 2024?
::: {.justify style="text-align: justify;"}
The results shows that Florida experienced the largest shift in vote margin toward Kamala Harris in the 2024 U.S. Presidential Election. The shift in margin was approximately 371,686 votes in her favor.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Ensure vote columns are numeric only if needed
if (!is.numeric(shapes_combined$donald_trump_republican_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(donald_trump_republican_2020 = parse_number(donald_trump_republican_2020))
}

if (!is.numeric(shapes_combined$donald_trump_republican_2024)) {
  shapes_combined <- shapes_combined |>
    mutate(donald_trump_republican_2024 = parse_number(donald_trump_republican_2024))
}

if (!is.numeric(shapes_combined$joe_biden_democratic_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(joe_biden_democratic_2020 = parse_number(joe_biden_democratic_2020))
}

if (!is.numeric(shapes_combined$kamala_harris_democratic_2024)) {
  shapes_combined <- shapes_combined |>
    mutate(kamala_harris_democratic_2024 = parse_number(kamala_harris_democratic_2024))
}


# Group by state and calculate margins
state_shift <- shapes_combined |>
  group_by(state) |>
  summarise(
    trump_2020 = sum(donald_trump_republican_2020, na.rm = TRUE),
    trump_2024 = sum(donald_trump_republican_2024, na.rm = TRUE),
    biden_2020 = sum(joe_biden_democratic_2020, na.rm = TRUE),
    harris_2024 = sum(kamala_harris_democratic_2024, na.rm = TRUE)
  ) |>
  mutate(
    margin_2020 = biden_2020 - trump_2020,
    margin_2024 = harris_2024 - trump_2024,
    shift_to_harris = margin_2024 - margin_2020
  )

# Identify the state with the largest shift toward Harris
top_state_shift <- state_shift |>
  arrange(desc(shift_to_harris)) |>
  slice_max(shift_to_harris, n = 1) |>
  mutate(
    state = str_to_title(state),
    shift_to_harris = comma(shift_to_harris)
  )

# Display result
top_state_shift |>
  st_drop_geometry() |>
  select(state, shift_to_harris) |>
  kable(
    caption = "State with the Largest Shift Toward Kamala Harris 2024",
    col.names = c("State", "Vote Margin Shift Toward Harris"),
    align = "lc",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 5.What is the largest county, by area, in this data set?
::: {.justify style="text-align: justify;"}
The table shows that Yukon-Koyukuk County in Alaska is the largest county by land area in the dataset, covering approximately 377,055,293,513 square meters. This is consistent with known geography, as Yukon-Koyukuk is widely recognized as the largest county-equivalent in the U.S. by land area
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Find the largest county by land area
largest_county_area <- shapes_combined |>
  filter(!is.na(aland)) |>
  arrange(desc(aland)) |>
  slice_max(aland, n = 1) |>
  select(state, county, aland)

# Display in clean table
largest_county_area |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    aland = scales::comma(aland)
  ) |>
  kable(
    caption = "Largest County by Land Area",
    col.names = c("State", "County", "Land Area (sq meters)"),
    align = "lcl",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 6.Which county has the highest voter density (voters per unit of area) in 2020?
::: {.justify style="text-align: justify;"}
The results reveal that in 2020 elections, New York County had the highest concentration of voters relative to its land size with a voter density of 0.01. This reflects the dense urban environment and high voter participation in a small geographic area.
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Ensure land area and total votes are numeric
if (!is.numeric(shapes_combined$aland)) {
  shapes_combined <- shapes_combined |>
    mutate(aland = parse_number(aland))
}
if (!is.numeric(shapes_combined$total_votes_cast_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(total_votes_cast_2020 = parse_number(total_votes_cast_2020))
}

# Calculate voter density (votes per square meter)
shapes_combined <- shapes_combined |>
  mutate(voter_density_2020 = total_votes_cast_2020 / aland)

# Identify county with highest voter density
highest_density <- shapes_combined |>
  filter(!is.na(voter_density_2020)) |>
  arrange(desc(voter_density_2020)) |>
  slice_max(voter_density_2020, n = 1) |>
  select(state, county, total_votes_cast_2020, aland, voter_density_2020)

# Display formatted table
highest_density |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    total_votes_cast_2020 = comma(total_votes_cast_2020),
    aland = comma(aland),
    voter_density_2020 = comma(voter_density_2020, accuracy = 0.01)
  ) |>
  kable(
    caption = "County with the Highest Voter Density in 2020",
    col.names = c("State", "County", "Total Votes (2020)", "Land Area (sq meters)", "Voter Density (votes/sq m)"),
    align = "lclrr",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```

### 7.Which county had the largest increase in voter turnout in 2024?
::: {.justify style="text-align: justify;"}
Between the 2020 and 2024 presidential elections, Montgomery County in Texas experienced the largest increase in voter turnout across all U.S. counties in the dataset. The total number of votes cast rose from 271,543 in 2020 to 307,258 in 2024, reflecting an increase of 35,715 votes. 
:::

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Convert total vote counts to numeric if not already
if (!is.numeric(shapes_combined$total_2020)) {
  shapes_combined <- shapes_combined |>
    mutate(total_2020 = parse_number(total_2020))
}

if (!is.numeric(shapes_combined$total_2024)) {
  shapes_combined <- shapes_combined |>
    mutate(total_2024 = parse_number(total_2024))
}

# Calculate change in turnout
shapes_combined <- shapes_combined |>
  mutate(turnout_change = total_2024 - total_2020)

# Identify county with largest increase
largest_turnout_increase <- shapes_combined |>
  filter(!is.na(turnout_change)) |>
  arrange(desc(turnout_change)) |>
  slice_max(turnout_change, n = 1) |>
  select(state, county, total_2020, total_2024, turnout_change)

# Display result in formatted table
largest_turnout_increase |>
  st_drop_geometry() |>
  mutate(
    state = str_to_title(state),
    county = str_to_title(county),
    total_2020 = comma(total_2020),
    total_2024 = comma(total_2024),
    turnout_change = comma(turnout_change)
  ) |>
  kable(
    caption = "County with the Largest Increase in Voter Turnout  2024",
    col.names = c("State", "County", "Votes in 2020", "Votes in 2024", "Increase in Turnout"),
    align = "lclrr",
    format = "html"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  )
```


